---
title: "introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  eval = FALSE
)
```

```{r setup}
library(course5assignment1)
```

The purpose of the __course5assignment1__ package is to clean up earthquake data 
from NOAA, to visualize a subset of the data on a timeline, and to visualize data 
in a map with annotations. 

## data clean-up 

The functions `eq_location_clean()` and `eq_clean_data()` can be used to clean up 
raw data from the NOAA website, and to generate a data.frame that can be supplied 
to the visualization functions. 

```{r}
# use `eq_location_clean()` to strip out the country name and converts names to title case
loc_name <- "JORDAN:  BAB-A-DARAA,AL-KARAK"
loc_name_cleaned <- eq_location_clean(loc_name) # "Bab-A-Daraa,Al-Karak"

# use `eq_clean_data` to obtain a data.frame that can be used for visualization.
data_raw <- readr::read_delim("./signif.txt") 
data_cleaned <- eq_clean_data(data_raw)

```

## Visualization with geom_timeline and geom_timeline_label 

`geom_timeline()` plots a time line of earthquakes, and `geom_timeline_label()` 
plots a time line and adds annotations to the earthquake data. `stat_timeline()` 
can be used to prepare data for geom_timeline or geom_timeline_label, or for 
other geom_* functions in __ggplot2__.

```{r}
# read data 
data_raw <- readr::read_delim("./signif.txt")
dat <- eq_clean_data(data_raw) %>%
  filter(COUNTRY %in% c("MEXICO", "ITALY"))

# use `stat_timeline` to prepare data 
# using default aesthetics 
ggplot(dat, aes(x=date)) +
   stat_timeline(geom="point")

# specifying aesthetics and other parameters 
ggplot(dat, aes(x=date, y=COUNTRY, size=INTENSITY, colour=TOTAL_DEATHS)) +
  stat_timeline(geom="point", x_min="1980-01-01", x_max="2017-01-01")

# use `geom_timeline` to plot a time line of earthquakes 
# using default aesthetics 
ggplot(dat, aes(x=date)) +
  geom_timeline()
# specifying aesthetics and other parameters  
ggplot(dat, aes(x=date, y=COUNTRY, colour=TOTAL_DEATHS, size=INTENSITY)) + 
  geom_timeline(alpha=.1, x_min="1980-01-01", x_max="2017-01-01")

# use `geom_timeline_label` to plot a time line of earthquakes and annotations
# using default aesthetics 
ggplot(dat, aes(x=date)) +
  geom_timeline_label()
# specifying aesthetics and other parameters
ggplot(dat, aes(x=date, y=COUNTRY, colour=TOTAL_DEATHS, size=INTENSITY)) +
  geom_timeline_label(alpha=.1, x_min="1980-01-01", x_max="2017-01-01", n_max=10)

```

## Visualization of earthquake data in a map with interactive annotations. 

The function `eq_map()` takes a data.frame of earthquake data and generates a 
leaflet object to visualize the data in a map. Annotations can be added 
interactively as pop-up text. `eq_create_label()` can be used to create complex
label for the map. 

```{r}
# use `eq_create_label` to create an HTML label for a data point
label_1 <- eq_create_label(dat[301,])  

# use `eq_map` to visualize earthquakes in a map with interactive annotations using 'date' 
readr::read_delim("./signif.txt") %>% 
  eq_clean_data() %>% 
  dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(date) >= 2000) %>% 
  eq_map(annot_col = "date")

# use `eq_map` to visualize earthquakes in a map with interactive annotations complex label created by `eq_create_label`
readr::read_delim("./signif.txt") %>% 
  eq_clean_data() %>% 
  dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(date) >= 2000) %>% 
  dplyr::mutate(popup_text = eq_create_label(.)) %>% 
  eq_map(annot_col = "popup_text")

```







